```{r function section, echo=FALSE,include=FALSE}

source("00.import.R")

```

---
title: "`r paste("Agonist-Induced Functional Analysis and Cell Sorting is a novel tool to uncover interneuron impairment: Fragile X Syndrome as a proof of concept")`"
author: "`r paste("K.Lebrigand")`"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    highlight: tango
    number_sections: yes
    theme: flatly
    toc: yes
---

# Integration of both datasets

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=6, fig.width=16}

wt <- readRDS("output/wt.baseline.rds")
wt.ampa <- readRDS("output/wt.ampa.clean.rds")
wt[['cluster']] <- wt@active.ident
wt.ampa[['cluster']] <- wt.ampa@active.ident
wt <- RenameCells(wt, add.cell.id = "baseline")
wt.ampa <- RenameCells(wt.ampa, add.cell.id = "ampa")

library(devtools)
library(ggplot2)
detach("package:Seurat", unload=TRUE)
library(Seurat, lib.loc = "/home/lebrigand/R-dev/")

features <- SelectIntegrationFeatures(object.list = c(wt,wt.ampa), nfeatures = 3000)
options(future.globals.maxSize = 4096 * 1024^2)
list <- PrepSCTIntegration(object.list = c(wt,wt.ampa), anchor.features = features, verbose = FALSE)
anchors <- FindIntegrationAnchors(object.list = list, normalization.method = "SCT", anchor.features = features, verbose = FALSE)
aggr <- IntegrateData(anchorset = anchors, normalization.method = "SCT", verbose = FALSE)
aggr <- RunPCA(aggr, verbose = FALSE)
aggr <- RunUMAP(aggr, dims = 1:30)
aggr <- FindNeighbors(object = aggr, do.plot=TRUE, dims = 1:30)
aggr <- FindClusters(object = aggr, resolution=0.6)

detach("package:Seurat", unload=TRUE)
library(Seurat)

#plots <- DimPlot(aggr, group.by = c("sample","cluster"), combine = FALSE)
#plots <- lapply(X = plots, FUN = function(x) x + theme(legend.position = "top") + guides(color = guide_legend(nrow = 3, byrow = TRUE, #override.aes = list(size = 3))))
#CombinePlots(plots)

DimPlot(aggr, label=TRUE)

```

# Correlation heatmap

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=6, fig.width=8}

mnmat <- c()
uniq <- unique(aggr@active.ident)
aggr@meta.data$cluster <- aggr@active.ident
for(i in 1:length(uniq)){
  mnmat <- cbind(mnmat, apply(as.matrix(aggr@assays$RNA@data[, aggr@meta.data$cluster==uniq[i]]), 1, mean))
}

colnames(mnmat) <- as.vector(unique(aggr@active.ident))
ct=cor(mnmat)
pheatmap(ct)

median(aggr@meta.data$nCount_RNA)
dim(aggr)
table(aggr@meta.data$sample)

```

## Gene Markers

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=10, fig.width=8 }

DefaultAssay(aggr) <- "RNA"
aggr <- NormalizeData(aggr)
aggr <- ScaleData(aggr)
markers <- FindAllMarkers(object = aggr, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
top5 <- markers %>% group_by(cluster) %>% top_n(5, avg_logFC)
DoHeatmap(aggr, features=top5$gene, size=3.5)
write.table(markers, file="output/integrated.clusters.csv", sep=",")

```

# Cluster relabeling

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=6, fig.width=8}

new.cluster.ids <- c("Inh-Sox4","Ex-Sept4","Ex-Top2a","Ex-Neurod2","Ex-Selm", "Inh-Synpr", "Inh-Top2a", "Sema3c", "Akap7")
names(x = new.cluster.ids) <- levels(x = aggr)
aggr <- RenameIdents(object = aggr, new.cluster.ids)

my_levels <- c("Ex-Neurod2","Ex-Sept4","Ex-Selm","Ex-Top2a","Inh-Top2a", "Inh-Sox4", "Inh-Synpr", "Sema3c", "Akap7")
aggr@active.ident <- factor(x = aggr@active.ident, levels = my_levels)

DimPlot(aggr, label=TRUE, cols = cell_type_color)

saveRDS(aggr, "output/aggr.rds")

freq_table <- table(aggr@active.ident, aggr@meta.data[, "sample"])
freq_table

pdf("figures.pdf", width=8, height=6, useDingbats=FALSE)
DimPlot(object = aggr, pt.size = 1, cols = cell_type_color)
DimPlot(object = aggr, group.by="sample", pt.size = 1, cols = cell_type_color)
FeaturePlot(aggr, c("Meis2","Neurod1","Neurod2","Gad2"), pt.size = 1, col=c("grey","red"))
FeaturePlot(aggr, c("Top2a","Top2a","Top2a","Top2a"), pt.size = 1, col=c("grey","red"))
dev.off()

pdf("barplot.pdf", width=6, height=8, useDingbats=FALSE)
freq_table <- prop.table(x = table(aggr@active.ident, aggr@meta.data[, "sample"]), margin = 2)*100
freq_table
tmp <- melt(freq_table)
names(tmp) <- c("cluster", "sample", "percentage")
#tmp<- tmp[seq(dim(tmp)[1],1),]
ggplot(tmp, aes(x=sample, y=percentage, fill=cluster)) + 
      geom_bar(stat="identity", position = position_fill(reverse = TRUE)) +
      scale_y_discrete(limits=c("Ex-Neurod2","Ex-Sept4","Ex-Selm","Ex-Top2a","Inh-Top2a", "Inh-Sox4", "Inh-Synpr", "Sema3c", "Akap7")) + 
      scale_fill_manual(values=cell_type_color) +
      theme_minimal()
dev.off()

DefaultAssay(aggr) <- "RNA"
aggr <- NormalizeData(aggr)
aggr <- ScaleData(aggr)
markers <- FindAllMarkers(object = aggr, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
write.table(markers, file="output/integrated.labels.csv", sep=",")

DefaultAssay(object = aggr) <- "RNA"
countTable <- data.frame(GetAssayData(aggr))
aggr[['cell_type']] <- aggr@active.ident
metadata <- aggr@meta.data
cell_type_order <- c("Ex-Neurod2","Ex-Sept4","Ex-Selm","Ex-Top2a","Inh-Top2a", "Inh-Sox4", "Inh-Synpr", "Sema3c", "Akap7")

x.markers <- read.table("output/integrated.labels.csv",sep = ",",header = T)
a <- x.markers %>%  filter(cluster%in%c("Ex-Neurod2","Ex-Sept4","Ex-Selm","Ex-Top2a","Inh-Top2a", "Inh-Sox4", "Inh-Synpr", "Sema3c", "Akap7"))%>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
markers <- as.character(unique(a[order(match(a$cluster, cell_type_order)),]$gene))
markers_flt_count <- countTable[markers, ]

# gene markers of manuscript --> heatmap2.pdf
#markers <- c('Dcx','Gad1','Gad2','Dlx6','Dlx1','Dlx5','Dlx2','Slc17a7','Slc17a6','Nrn1','Neurod1','Neurod2','Top2a','Ube2c','Mki67','Hmgb2','Cenpf','Selm','Meis1','Pax6','E2f1','Cog7','Cd63','Tiam2','Nrxn3','Pbx1','Sox4','Synpr','Calb2','Pbx3','Sema3c','Id2','Cnr1','Cldn3','Akap7','Cttnbp2','Hes6','Ccnd2','Apc','Nrxn1','Rbfox3','Map1b','Snap25','Grin2b','Pclo','Mapt','Kcnb2','Kcna1')
#markers_flt_count <- countTable[markers, ]

markers_count <- matrix(0, nrow = length(markers), ncol = length(cell_type_order))
colnames(markers_count) <- c("Ex-Neurod2","Ex-Sept4","Ex-Selm","Ex-Top2a","Inh-Top2a", "Inh-Sox4", "Inh-Synpr", "Sema3c", "Akap7")
rownames(markers_count) <- markers

for (i in 1:ncol(markers_count)){
  print(colnames(markers_count)[i])
  cell_names <- rownames(metadata[metadata$cell_type == colnames(markers_count)[i], ])
  markers_count[, colnames(markers_count)[i]] <- rowMeans(markers_flt_count[, cell_names])
}

flexible_normalization <- function(data_in, by_row=TRUE){
  if(by_row){
    row_mean <- apply(data_in,1,mean)
    row_sd   <- apply(data_in,1,sd)
    output <- data_in
    for(i in 1:dim(data_in)[1]){
      output[i,] <- (data_in[i,] - row_mean[i])/row_sd[i]
    }
  }
  #### if by column
  if(!by_row){
    col_mean <- apply(data_in,2,mean)
    col_sd   <- apply(data_in,2,sd)
    output <- data_in
    for(i in 1:dim(data_in)[2]){
      output[,i] <- (data_in[,i] - col_mean[i])/col_sd[i]
    }
  }
  return(output)
}

markers_count <- flexible_normalization(markers_count, by_row = T)
markers_count[markers_count > 2] <- 2
markers_count[markers_count < -1.5] <- -1.5

annotation <- data.frame(CellType = factor(colnames(markers_count), levels = cell_type_order))
rownames(annotation) <- colnames(markers_count)
ann_colors <- list(CellType = cell_type_color)
pp <- pheatmap::pheatmap(markers_count, cluster_rows = F, cluster_cols = F, 
                         annotation_col = annotation, annotation_colors = ann_colors,     
                         angle_col = 0, annotation_legend = F)
pdf("figures/heatmap.pdf", width=8, height=10, useDingbats=FALSE)
pp
dev.off()


```

# Session Info

```{r sessinf}
sessionInfo()
```

